Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2019-04-02 paper.tex

\algblock{Input}{EndInput}
\algnotext{EndInput}
\algblock{Output}{EndOutput}
\algnotext{EndOutput}
\newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}

\begin{document}

\title{Performance Evaluation of Big Data Processing Strategies for Neuroimaging}

@@

\begin{abstract}
    Neuroimaging datasets are rapidly growing in size as a result of 
    advancements in image acquisition methods, open-science and data sharing. 
    However, current neuroimaging data engines have not adopted any strategies
    to address the rise
    in neuroimaging data. Here, we 
    evaluate three Big Data processing strategies (in-memory computing, 
    data locality and lazy evaluation) on typical neuroimaging use 
    cases, represented by the BigBrain dataset. We contrast these various 
    strategies using Apache Spark and Nipype as our 
    representative Big Data and neuroimaging processing engines, on Dell EMC's 
    Top-500 cluster. 
    Big Data thresholds were modeled by comparing the data-write rate of the 
    application to the filesystem-application concurrent write rate. 
    This model acknowledges the 
fact that page caching provided by the Linux kernel is critical to the 
performance of Big Data applications. Results show that in-memory 
computing alone speeds-up executions by a factor of up to 1.6, whereas 
when combined with data locality, this factor reaches 5.3. Lazy evaluation
strategies
were found to increase the likelihood of cache hits, further improving
    processing time. Such important 
speed-up values are likely to be observed on typical image processing 
operations performed on images of size larger than 75GB. 
We conclude that Big Data processing strategies are 
worth developing for neuroimaging 
applications. 
\end{abstract}

@@

In Big Data applications, there is a significant cost associated with data movement.
Big Data processing engines have addressed these issues by diminishing the amount 
of data movement that occurs during the execution of an application. The 
MapReduce~\cite{dean2008mapreduce} framework popularized the concept of locality-aware
scheduling, which ensured tasks were scheduled nearest to where the data the data was
located (data locality), thereby reducing the number of network transfers in an application.
However, it was found that performance in MapReduce could be improved upon by also
reducing the disk I/O between tasks by keeping data in-memory whenever possible,
and thus, in-memory computing was added to the Apache Spark~\cite{zaharia2016apache} framework.
Frameworks like Apache Spark also included the lazy evaluation strategy, which performs computations
only when invoked. This allows for additional optimizations such as the regrouping of tasks
and only invoking what is absolutely necessary to obtain the desired result.

Frameworks such as
MapReduce and Spark have become mainstream tools for data analytics, however, many
others, such as Dask~\cite{rocklin2015dask}, are emerging.
Meanwhile, several scientific 
domains including bioinformatics, physics or astronomy, have entered 
the Big Data era due to increasing data volumes and variety. 
Nevertheless, the adoption of Big Data engines for scientific data analysis 
remains limited, perhaps due to the widespread availability of 
scientific processing engines such as Pegasus~\cite{deelman2005pegasus} or
Taverna~\cite{oinn2004taverna}, and the adaptations required in Big 
Data processing engines for scientific computing.

@@

Infrastructure-wise, we focus on the case of High-Performance Computing 
(HPC) clusters that are typically available through University 
facilities or national computing infrastructures such as 
\href{xsede.org}{XSEDE}, \href{http://computecanada.ca}{Compute Canada} 
or \href{http://www.prace-ri.eu}{PRACE}, as neuroscientists typically use such platforms.
 We assume that HPC systems are 
multi-tenant, that compute nodes are accessible through a batch 
scheduler, and that a file system shared among the compute nodes is 
available. We intentionally did not consider distributed, 
HDFS-like file systems, as initiatives to deploy them in HPC 
centers, for instance Hadoop on-demand~\cite{krishnan2011myhadoop}, have not 
become mainstream yet.

@@

Although Spark uses in-memory computing, it is not necessary for all the data to
fit in memory. Spark will spill any RDD elements that do not fit in memory to disk.
Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur within a single application, Spark implements a
Least-Recently Used (LRU) eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. As demonstrated in~\cite{freeman2014mapping}, caching
significantly improves processing times of iterative algorithms where 
RDDs are reused. It can be of even greater importance if the RDD is costly to 
recompute.

@@

Data storage location is critical to the performance of Big Data 
applications on HPC clusters.
Data may reside in the engine memory, on 
a file system whose contents reside in virtual memory (for instance 
tmpfs), on disks local to the processing node, or on a shared 
file system. Table~\ref{table:features} summarizes the Big Data 
strategies that can be used depending on the data location. In 
addition, lazy evaluation is available in Spark regardless of data 
location. The remainder of this Section explains this Table and 
provides related performance models.
\begin{table}
\centering
\begin{tabular}{c|cc}
   \rowcolor{headcolor}
    Data Location                 & In-Memory     & Data Locality        \\
    \rowcolor{headcolor}
                                  & Computing     &                     \\
                                  \hline          
In-memory                         & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                      \\
tmpfs                             & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                  \\
Local Disk                        & \cellcolor{orange!25}Page Caching  & \cellcolor{green!25}Yes                  \\
Shared File System                & \cellcolor{orange!25}Page Caching  & \cellcolor{red!25}No                
\end{tabular}
\caption{Big Data strategies on a shared HPC cluster.}
\label{table:features}
\end{table}

@@

 The main difference between storing the data in the engine memory, as in Spark, 
 and simply writing to an in-memory file system, such as tmpfs, is 
 what happens when the processed data fills up the available 
 memory. When engine memory is used, the engine must cleanup 
 unused data to avoid crashes. When an in-memory 
 file system is used, the user is responsible ensuring that the filesystem does not
 reach capacity. Should it be necessary to free up additional memory, the kernel will swap filesystem memory to disk,
 and the performance will become that of local disk writes. In our experiments, 
 we will explore the configuration where the data consumed by the 
 application approaches the threshold of available memory.

@@

In general, an application should ensure that $\dot d$ remains negative 
or null, leading to the following inequality:
\begin{equation}
\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
\end{equation}
This defines a data-write rate beyond which the page cache becomes 
asymptotically useless. It should be noted that the transient phase 
during which the page cache fills up might last a significant amount of time, 
in particular when $\dot d$ is positive and small. We intentionally do not model the 
transient phase as it requires detailed knowledge of difficult to estimate parameters  
such as the page cache size and the initial amount 
of data in it ($d_0$).

@@

 All experiments were executed on 
 \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell 
 EMC's Zenith cluster}, a Top-500 machine in the 
 \href{https://www.dellemc.com/en-us/solutions/high-performance-computing/HPC-AI-Innovation-Lab.htm}{Dell 
 EMC HPC and AI Innovation Lab}, running Slurm. For the Spark 
 experiments, a Spark cluster was started on a
 Slurm allocation comprised of 16 dedicated nodes.
 Each Compute node has Red Hat Enterprise Linux Server release 7.4 (Maipo) 
 as the base operating system with kernel version 3.10.0-693.17.1.el7.x86\_64
 (patched for Spectre/Meltdown). Dell EMC PowerEdge C6420 with dual Intel Xeon
 Gold 6148/F processors (40 cores per node) and 192GB ($12\times16$~GB), 2666 MHz
 memory, serve as the compute nodes. Each compute has a 120GB M.2 SATA SSD as 
 local disk. A Dell HPC Lustre Solution with a raw storage of 960TB is 
 accesible on each compute node through a 100~Gb/s Intel OmniPath network. All
 the nodes connect to a director switch in a 1:1 non-blocking topology.
 The realized write bandwidth of the local disk, 
 Lustre file system and tmpfs were measured by sequentially writing various numbers
 of image blocks containing random intensities, to avoid 
 caching effects (measure\_bandwidth.py script).
 They are reported in Table~\ref{table:bdwdths}.

\begin{table}
\centering
\begin{tabular}{c|c}
\rowcolor{headcolor}
Data location & Measured write bandwidths (MB/s)\\
\hline
tmpfs                 & 1377.18 \\
Local disk ($\delta$) & 193.64  \\
Lustre   ($\Delta$)   & 504.03 \\
\end{tabular}
\caption{Measured bandwidths}
\label{table:bdwdths}
\end{table}

@@

Processing large images is only considered to be part of the Big Data problem 
in neuroscience. The other being the processing large MRI datasets, 
that is, datasets consisting of many small brain images belonging to 
various different subjects. This situation is commonly observed in 
functional MRI (fMRI), where it is becoming increasingly common to 
process data from hundreds of subjects. Although we have not explored 
explicitly the processing of large MRI datasets, the 75GB BigBrain is 
within the size ballpark~\cite{van2014human} of 
MRI datasets commonly processed in today's studies.

@@

\hline
  \multicolumn{10}{c}{}\\ 
 \rowcolor{headcolor}
  \multicolumn{10}{c}{Experiment 4: Image Size}\\
  \hline
  \rowcolor{headcolor}
 image  &
 D (GB) & C (s) & D/C (MB/s) &
 $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
(D/C)/($\delta/\gamma$) & (D/C)/($\Delta/\Gamma$)\\
 \hline
 BigBrain      & 750   & 2,200   & 349.1     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} 16.2 & \cellcolor{red!20} 87.3   \\
 Half BigBrain & 375   & 2,200   & 174.6     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} 8.1 & \cellcolor{red!20} 43.7   \\
 MRI           & 0.127 & 2,200   & 0.06      & 9  & 21.5  & 125 & 4.0 &  \cellcolor{green!20} 0.003 & \cellcolor{green!20} 0.015  
\end{tabular}
\caption{Experiment conditions. Red cells denote the conditions where 
the inequalities in Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs} do not hold, i.e., the page cache is 
asymptotically useless. Green cells show the conditions where the page cache covers all data writes.}
\label{table:experiments}
\end{table*}

@@

As a third incrementation experiment, we were interested in the effects of 
chunk size on Big Data strategies. Naturally, a greater chunk size signifies 
a decrease in parallelism. However, it also signifies an increase in 
sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
this experiment we partitioned the complete BigBrain image into 30, 125 
and 750 chunks, corresponding to $\gamma$ values of 2, 9 and 25 
respectively. While Spark attempted to load-balance the data, it used 
up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
use up as many cores as possible. Unlike the previous 
experiment, the data-write rate was kept static at 178.6MB/s, however, this 
ratio ensured that different regions of the inequality were reached 
depending on amount of parallelism. The number of iterations was
fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.

For our fourth and final incrementation experiment, we investigated the effects 
of the strategies on different image sizes. We selected the 75GB BigBrain, the
38GB half BigBrain and the 13M T1W MRI image for this experiment. The number of
chunks was fixed to 125. Similarly to the previous experiment, the 
total sequential compute time was fixed (10 iterations, 1.76 seconds 
per task), however, due to varying size in total data processed (D), 
the data-write rate varied. Once again, we ensured that the data-write rate fell in 
multiple different regions of the inequality. The data-write rate ranged from 
349.1 MB/s for BigBrain and 174.6MB/s for half BigBrain, to 0.06MB/s for the MRI 
image. Only the 0.06MB/s MRI satisfied the inequality for both Lustre and local 
disk.

@@

\begin{figure*}
\begin{subfigure}{\columnwidth}
    \centering
    \begin{subfigure}{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{results/figures/s-iterations.pdf}%
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{results/figures/s-iterations.pdf}%
    \end{subfigure}
    \includegraphics[width=\linewidth]{results/figures/iterations.pdf}
    \caption{Experiment 1: complete BigBrain, 125 chunks, 3.44-second tasks.}\label{fig:iterations}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/cputime.pdf}
    \caption{Experiment 2: complete BigBrain, 125 chunks, 10 iterations.}\label{fig:cputime}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/numchunks.pdf}
    \caption{Experiment 3: complete BigBrain, 10 iterations, C= 4,400s.}\label{fig:numchunks}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/datasize.pdf}
    \caption{Experiment 4: 125 chunks, 10 iterations, 1.76-second tasks.}\label{fig:datasize}
\end{subfigure}
\caption{Experiment results: Makespans of Spark and Nipype writing to memory, tmpfs, local disk and Lustre.}
\label{fig:results}
\end{figure*}
\subsection{Experiment 1: Number of Iterations}

@@

\subsection{Page Cache Model Evaluations}
\begin{figure*}
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/local-incrementation.pdf}
\caption{Local Disk}
\label{fig:modeleval-local}
    \end{subfigure}%
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/lustre-incrementation.pdf}
\caption{Lustre}
\label{fig:modeleval-lustre}
\end{subfigure}
    \caption{Page cache model evaluation. Grey 
             regions denote areas that violate model predictions.}
\label{fig:modeleval}
\end{figure*}

In order to evaluate the page cache model, we compared the observed 
speedup ratio provided by in-memory computing to the (D/C) / 
($\delta$/$\gamma$) and (D/C) / ($\Delta$/$\Gamma$) ratios 
(Fig.~\ref{fig:modeleval}). Speed-up ratios were computed as the ratio 
between the makespan obtained with Spark on local disk or Lustre, and 
the makespan obtained with Spark for in-memory computing. Experiments 
for which there was no in-memory equivalent (i.e. BigBrain split into 
30 
chunks) 
were not considered.

Results show that, overall, the model correctly predicted the effect of 
page cache on processing times for local disk and Lustre. That is, the 
speed-up provided by in-memory computing was larger than 1 for data-write 
rates larger than $\delta/\gamma$ (local disk) or $\Delta/\Gamma$ 
(Lustre). Conversely, the speed-up provided by in-memory computing 
remained close to 1 for data-write rates smaller than $\delta/\gamma$ (local 
disk) or $\Delta/\Gamma$ (Lustre). The two points close to the origin 
correspond to the sequential processing of the MRI image by Spark 
mentioned previously.

@@

The results also allow us to speculate on the effect of in-memory 
computing on the pre-processing of functional MRI, another typical use 
case in neuroimaging. Assuming an average processing time of 20 minutes 
per subject, which is a ballpark value commonly observed with the 
popular \href{https://www.fil.ion.ucl.ac.uk/spm/}{SPM} or \href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki/}{FSL} packages,
an input data size of 100MB per subject, 
and an output data size of 2GB (20-fold increase compared to input 
size), the data-write rate would be 1.8MB/s, which would reach the 
$\delta/\gamma$ threshold measured on this cluster for $\gamma=108$, 
that is, if 108 subjects were processed on the same node. This is very 
unlikely as the number of CPUs per node was 40. We therefore conclude 
that in-memory computing is likely to be useless for fMRI analysis.
Naturally, this estimate is strongly dependent on the characteristics of the cluster.

@@

As for the fMRI speculation, the data-write 
rate of 1.8MB/s would reach the $\Delta/\Gamma$ threshold for 
$\Gamma=280$, which is a realistic number of subjects to process on a 
complete cluster. Naturally, this estimate is highly dependent on the observed bandwidth
of the shared file system ($\Delta$).

@@

A common recommendation in Spark is to limit the number of cores per 
executors to 5, to preserve a good I/O 
throughput\footnote{\url{http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2}}, 
but the rationale for this recommendation is hardly explained. We 
believe that throughput degradation observed with more than 5 cores per 
executor might be coming from full page caches.

@@

It was expected that Spark would experience longer processing times, particularly
with the small datasets, due to Java serializing. This was not found to be the case. 
Unlike Spark, Nipype performs a more thorough provenance capture, potentially owing to longer
processing times.

@@

Big Data performance optimization strategies help improve performance of typical 
neuroimaging applications. Our experiments indicate that overall, in-memory computing 
enables greater speedups than what can be obtained by using page cache 
and tmpfs. While page cache and tmpfs do give memory-like performance, 
they are likely to fill up faster than memory, leading to increased 
performance penalties when compared to in-memory computing. We conclude 
that extending Big Data processing engines to better support 
neuroimaging applications, including developing their provenance, 
fault-tolerance, and reproducibility features, is a worthwhile project.

>>>>

\algblock{Input}{EndInput}
\algnotext{EndInput}
\algblock{Output}{EndOutput}
\algnotext{EndOutput}
\newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
\begin{document}

\title{Performance Evaluation of Big Data Processing Strategies for Neuroimaging}
\pagenumbering{gobble}

@@

\begin{abstract}
    Neuroimaging datasets are rapidly growing in size as a result of 
    advancements in image acquisition methods, open-science and data sharing. 
    However, the adoption of Big Data processing strategies by neuroimaging
    processing engines remains limited. Here, we 
    evaluate three Big Data processing strategies (in-memory computing, 
    data locality and lazy evaluation) on typical neuroimaging use 
    cases, represented by the BigBrain dataset. We contrast these various 
    strategies using Apache Spark and Nipype as our 
    representative Big Data and neuroimaging processing engines, on Dell EMC's 
    Top-500 cluster. 
    Big Data thresholds were modeled by comparing the data-write rate of the 
    application to the filesystem bandwidth and number of concurrent processes. 
    This model acknowledges the 
fact that page caching provided by the Linux kernel is critical to the 
performance of Big Data applications. Results show that in-memory 
computing alone speeds-up executions by a factor of up to 1.6, whereas 
when combined with data locality, this factor reaches 5.3. Lazy evaluation
strategies
were found to increase the likelihood of cache hits, further improving
    processing time. Such important 
speed-up values are likely to be observed on typical image processing 
operations performed on images of size larger than 75GB. A ballpark 
speculation from our model showed that in-memory computing alone will
not speed-up current 
functional MRI analyses unless coupled with data 
locality and processing around 280 subjects concurrently. Furthermore, we observe
    that emulating in-memory computing 
using in-memory file systems (tmpfs) does not reach the performance of an in-memory engine,
    presumably due to swapping to disk and the lack of data cleanup.
We conclude that Big Data processing strategies are 
worth developing for neuroimaging 
applications. 
\end{abstract}

@@

Big Data processing engines have significantly improved the performance of Big Data
applications by diminishing the amount 
of data movement that occurs during the execution of an application. Locality-aware
scheduling, introduced by the MapReduce~\cite{dean2008mapreduce} framework, reduced the overall
costs associated to network transfer of data by scheduling tasks to the nodes located nearest to
the data. Reduction of data movement was further improved upon through in-memory computing~\cite{zaharia2016apache},
which ensures that data is maintained in memory between tasks whenever possible. To further
reduce the cost of data movement, lazy evaluation, the process of performing computations only when invoked,
was leveraged by Big Data frameworks to enable further optimizations such as regrouping of tasks and
computing only what is necessary.

Frameworks such as
MapReduce and Spark~\cite{zaharia2016apache} have become mainstream tools for data analytics, although many
others, such as Dask~\cite{rocklin2015dask}, are emerging.
Meanwhile, several scientific 
domains including bioinformatics, physics or astronomy, have entered 
the Big Data era due to increasing data volumes and variety. 
Nevertheless, the adoption of Big Data engines for scientific data analysis 
remains limited, perhaps due to the widespread availability of 
scientific processing engines such as Pegasus~\cite{deelman2005pegasus} or
Taverna~\cite{oinn2004taverna}, and the adaptations required in Big 
Data processing engines for scientific computing.

@@

In terms of infrastructure, we focus on the case of High-Performance Computing 
(HPC) clusters that are typically available through University 
facilities or national computing infrastructures such as 
\href{xsede.org}{XSEDE}, \href{http://computecanada.ca}{Compute Canada} 
or \href{http://www.prace-ri.eu}{PRACE}, as neuroscientists typically use such platforms.
 We assume that HPC systems are 
multi-tenant, that compute nodes are accessible through a batch 
scheduler, and that a file system shared among the compute nodes is 
available. We intentionally did not consider distributed, 
HDFS-like file systems, as initiatives to deploy them in HPC 
centers, for instance Hadoop on-demand~\cite{krishnan2011myhadoop}, have not 
become mainstream yet.

@@

Although Spark uses in-memory computing, it is not necessary for all the data to
fit in memory. Spark will spill any RDD elements that cannot be maintained in memory to disk.
Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur within a single application, Spark implements a
Least-Recently Used (LRU) eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. As demonstrated in~\cite{freeman2014mapping}, caching
significantly improves processing times of iterative algorithms where 
RDDs are reused. It can be of even greater importance if the RDD is costly to 
recompute.

@@

Data storage location is critical to the performance of Big Data 
applications on HPC clusters.
Data may reside in the engine memory, on 
a file system whose contents reside in virtual memory (for instance 
tmpfs), on disks local to the processing node, or on a shared 
file system. Table~\ref{table:features} summarizes the Big Data 
strategies that can be used depending on the data location. In 
addition, lazy evaluation is available in Spark regardless of data 
location. The remainder of this Section explains this Table and 
provides related performance models.
\begin{table}
\centering
\begin{tabular}{c|cc}
   \rowcolor{headcolor}
    Data Location                 & In-Memory     & Data Locality        \\
    \rowcolor{headcolor}
                                  & Computing     &                     \\
                                  \hline          
In-memory                         & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                      \\
tmpfs                             & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                  \\
Local Disk                        & \cellcolor{orange!25}Page Caching  & \cellcolor{green!25}Yes                  \\
Shared File System                & \cellcolor{orange!25}Page Caching  & \cellcolor{red!25}No                
\end{tabular}
\setlength{\belowcaptionskip}{-10pt}
\caption{Big Data strategies on a shared HPC cluster.}
\label{table:features}
\end{table}

@@

 The main difference between storing the data in the engine memory, as in Spark, 
 and simply writing to an in-memory file system, such as tmpfs, is 
 what happens when the processed data fills up the available 
 memory. When engine memory is used, the engine must cleanup 
 unused data to avoid crashes. When an in-memory 
 file system is used, the user is responsible ensuring that the filesystem does not
 reach capacity. Should it be necessary to free up additional memory, the kernel will swap filesystem memory to disk
 and the performance will become that of local disk writes. In our experiments, 
 we will explore the configuration where the data consumed by the 
 application approaches the threshold of available memory.

@@

In general, an application should ensure that $\dot d$ remains negative 
or null, leading to the following inequality:
\begin{equation}
\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
\end{equation}
This defines a D/C (data-write) rate beyond which the page cache becomes 
asymptotically useless. It should be noted that the transient phase 
during which the page cache fills up might last a significant amount of time, 
in particular when $\dot d$ is positive and small. We intentionally do not model the 
transient phase as it requires detailed knowledge of difficult to estimate parameters  
such as the page cache size and the initial amount 
of data in it ($d_0$).

@@

 All experiments were executed on 
 \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell 
 EMC's Zenith cluster}, a Top-500 machine in the 
 \href{https://www.dellemc.com/en-us/solutions/high-performance-computing/HPC-AI-Innovation-Lab.htm}{Dell 
 EMC HPC and AI Innovation Lab}, running Slurm. For the Spark 
 experiments, a Spark cluster was started on a
 Slurm allocation comprised of 16 dedicated nodes.
 Each Compute node has Red Hat Enterprise Linux Server release 7.4 (Maipo) 
 as the base operating system with kernel version 3.10.0-693.17.1.el7.x86\_64
 (patched for Spectre/Meltdown). Dell EMC PowerEdge C6420 with dual Intel Xeon
 Gold 6148/F processors (40 cores per node) and 192GB ($12\times16$~GB), 2666 MHz
 memory, serve as the compute nodes. Each compute has a 120GB M.2 SATA SSD as 
 local disk. A Dell HPC Lustre Solution with a raw storage of 960TB is 
 accesible on each compute node through a 100~Gb/s Intel OmniPath network. All
 the nodes connect to a director switch in a 1:1 non-blocking topology.
 The realized write bandwidth of the local disk, 
 Lustre file system and tmpfs were measured by sequentially writing various numbers
 of image blocks containing random intensities, to avoid 
 caching effects (see: \href{https://github.com/big-data-lab-team/paper-in-mem-locality/blob/master/benchmark_scripts/measure_bandwidth.py}{measure\_bandwidth.py}).
 They are reported in Table~\ref{table:bdwdths}.

\begin{table}
\centering
\begin{tabular}{c|c}
\rowcolor{headcolor}
Data location & Measured write bandwidths (MB/s)\\
\hline
tmpfs                 & 1377.18 \\
Local disk ($\delta$) & 193.64  \\
Lustre   ($\Delta$)   & 504.03 \\
\end{tabular}
\setlength{\belowcaptionskip}{-10pt}
\caption{Measured bandwidths}
\label{table:bdwdths}
\end{table}

@@

Processing large images is only considered to be part of the Big Data problem 
in neuroscience. The other problem being the processing large MRI datasets, 
that is, datasets consisting of many small brain images belonging to 
various different subjects. This situation is commonly observed in 
functional MRI (fMRI), where it is becoming increasingly common to 
process data from hundreds of subjects. Although we have not explored 
explicitly the processing of large MRI datasets, the 75GB BigBrain is 
within the size ballpark~\cite{van2014human} of 
MRI datasets commonly processed in today's studies.

@@

\hline
  \multicolumn{10}{c}{}\\ 
 \rowcolor{headcolor}
  \multicolumn{10}{c}{Experiment 4: Image Size}\\
  \hline
  \rowcolor{headcolor}
 image  &
 D (GB) & C (s) & D/C (MB/s) &
 $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
(D/C)/($\delta/\gamma$) & (D/C)/($\Delta/\Gamma$)\\
 \hline
 BigBrain      & 750   & 2,200   & 349.1     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} 16.2 & \cellcolor{red!20} 87.3   \\
 Half BigBrain & 375   & 2,200   & 174.6     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} 8.1 & \cellcolor{red!20} 43.7   \\
 MRI           & 0.127 & 2,200   & 0.06      & 9  & 21.5  & 125 & 4.0 &  \cellcolor{green!20} 0.003 & \cellcolor{green!20} 0.015  
\end{tabular}
\setlength{\belowcaptionskip}{-10pt}
\caption{Experiment conditions. Red cells denote the conditions where 
the inequalities in Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs} do not hold, i.e., the page cache is 
asymptotically useless. Green cells show the conditions where the page cache covers all data writes.}
\label{table:experiments}
\end{table*}

@@

As a third incrementation experiment, we were interested in the effects of 
chunk size on Big Data strategies. Naturally, a greater chunk size signifies 
a decrease in parallelism. However, it also signifies an increase in 
sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
this experiment we partitioned the complete BigBrain image into 30, 125 
and 750 chunks, corresponding to $\gamma$ values of 2, 9 and 25 
respectively. While Spark attempted to load-balance the data, it used 
up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
use up as many cores as possible. Unlike the previous 
experiment, the D/C rate was kept static at 178.6MB/s, however, this 
ratio ensured that different regions of the inequality were reached 
depending on amount of parallelism. The number of iterations was
fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.

For our fourth and final incrementation experiment, we investigated the effects 
of the strategies on different image sizes. We selected the 75GB BigBrain, the
38GB half BigBrain and the 13M T1W MRI image for this experiment. The number of
chunks was fixed to 125. Similarly to the previous experiment, the 
total sequential compute time was fixed (10 iterations, 1.76 seconds 
per task), however, due to varying size in total data processed (D), 
the D/C rate varied. Once again, we ensured that the D/C rate fell in 
multiple different regions of the inequality. The D/C rate ranged from 
349.1 MB/s for BigBrain and 174.6MB/s for half BigBrain, to 0.06MB/s for the MRI 
image. Only the 0.06MB/s MRI satisfied the inequality for both Lustre and local 
disk.

@@

\begin{figure*}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{results/figures/iterations.pdf}%
    \caption{Experiment 1: complete BigBrain, 125 chunks, 3.44-second tasks.}\label{fig:iterations}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/cputime.pdf}
    \caption{Experiment 2: complete BigBrain, 125 chunks, 10 iterations.}\label{fig:cputime}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/numchunks.pdf}
    \caption{Experiment 3: complete BigBrain, 10 iterations, C= 4,400s.}\label{fig:numchunks}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/datasize.pdf}
    \caption{Experiment 4: 125 chunks, 10 iterations, 1.76-second tasks.}\label{fig:datasize}
\end{subfigure}
\setlength{\belowcaptionskip}{-10pt}
\caption{Experiment results: Makespans of Spark and Nipype writing to memory, tmpfs, local disk and Lustre.}
\label{fig:results}
\end{figure*}
\subsection{Experiment 1: Number of Iterations}

@@

\subsection{Page Cache Model Evaluations}
\begin{figure*}
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/local-incrementation.pdf}
\caption{Local Disk}
\label{fig:modeleval-local}
    \end{subfigure}%
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/lustre-incrementation.pdf}
\caption{Lustre}
\label{fig:modeleval-lustre}
\end{subfigure}
\setlength{\belowcaptionskip}{-10pt}
    \caption{Page cache model evaluation. Grey 
             regions denote areas that violate model predictions.}
\label{fig:modeleval}
\end{figure*}

In order to evaluate the page cache model, we compared the observed 
speedup ratio provided by in-memory computing to the (D/C) / 
($\delta$/$\gamma$) and (D/C) / ($\Delta$/$\Gamma$) ratios 
(Fig.~\ref{fig:modeleval}). Speed-up ratios were computed as the ratio 
between the makespan obtained with Spark on local disk or Lustre, and 
the makespan obtained with Spark for in-memory computing. Experiments 
for which there were no in-memory equivalent (i.e. BigBrain split into 
30 
chunks) 
were not considered.

Results show that, overall, the model correctly predicted the effect of 
page cache on processing times for local disk and Lustre. That is, the 
speed-up provided by in-memory computing was larger than 1 for D/C 
rates larger than $\delta/\gamma$ (local disk) or $\Delta/\Gamma$ 
(Lustre). Conversely, the speed-up provided by in-memory computing 
remained close to 1 for D/C rates smaller than $\delta/\gamma$ (local 
disk) or $\Delta/\Gamma$ (Lustre). The two points close to the origin 
correspond to the sequential processing of the MRI image by Spark 
mentioned previously.

@@

The results also allow us to speculate on the effect of in-memory 
computing on the pre-processing of functional MRI, another typical use 
case in neuroimaging. Assuming an average processing time of 20 minutes 
per subject, which is a ballpark value commonly observed with the 
popular \href{https://www.fil.ion.ucl.ac.uk/spm/}{SPM} or \href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki/}{FSL} packages,
an input data size of 100MB per subject, 
and an output data size of 2GB (20-fold increase compared to input 
size), the D/C rate would be 1.8MB/s, which would reach the 
$\delta/\gamma$ threshold measured on this cluster for $\gamma=108$, 
that is, if 108 subjects were processed on the same node. This is very 
unlikely as the number of CPUs per node was 40. We therefore conclude 
that in-memory computing is likely to be useless for fMRI analysis.
Naturally, this estimate is strongly dependent on the characteristics of the cluster.

@@

As for the fMRI speculation, the D/C 
rate of 1.8MB/s would reach the $\Delta/\Gamma$ threshold for 
$\Gamma=280$, which is a realistic number of subjects to process on a 
complete cluster. Naturally, this estimate is highly dependent on the observed bandwidth
of the shared file system ($\Delta$).

@@

A common recommendation in Spark is to limit the number of cores per 
executors to 5, to preserve a good I/O 
throughput (see \href{http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2}{Cloudera blog}), 
but the rationale for this recommendation is hardly explained. We 
believe that throughput degradation observed with more than 5 cores per 
executor might be coming from full page caches.

@@

It was expected that Spark would experience longer processing times, particularly
with the small datasets, due to Java serialization. This was not found to be the case. 
Unlike Spark, Nipype performs a more thorough provenance capture, potentially owing to longer
processing times.

In this paper, we analyzed the effects of Big Data strategies on an map-only artificial
neuroimaging pipeline. This allowed us to examine the effects of these strategies
without being significantly obscured by other conditions. Studying the effects 
of such strategies on map-reduce type
workflows, in addition to real neuroimaging pipelines, would allows us to gain
further insight on the added
value of the Big Data performance stragies for neuroimaging use cases remains to be
done.

@@

Big Data performance optimization strategies help improve performance of typical 
neuroimaging applications. Our experiments indicate that overall, in-memory computing 
enables greater speedups than what can be obtained by using page cache 
and tmpfs. While page cache and tmpfs do give memory-like performance, 
they are likely to fill up faster than memory, leading to increased 
performance penalties when compared to in-memory computing. We conclude 
that extending Big Data processing engines to better support 
neuroimaging applications, including developing their provenance, 
fault-tolerance, and reproducibility features, is worthwhile.

Data locality plays an important role in application 
performance. Local disk was found to perform better than the shared 
filesystem despite having lower bandwidth, due to increased contention on 
the shared filesystem. Since local disk typically has less 
contention than shared filesystems, it is recommended to store data locally.
However, using local storage without a distributed file system 
may limit fault tolerance.
